---
title: "Activity Monitoring Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/valel/Downloads/PA")
file_path <- "activity.csv"
data <- read.csv(file_path, stringsAsFactors = FALSE)
str(data)
```{r steps-per-day}
library(dplyr)
steps_per_day <- data %>%
  group_by(date) %>%
  summarize(total_steps = sum(steps, na.rm = TRUE))
hist(steps_per_day$total_steps, breaks = 20, main = "Total Steps per Day", xlab = "Total Steps", ylab = "Frequency")
mean_steps <- mean(steps_per_day$total_steps, na.rm = TRUE)
median_steps <- median(steps_per_day$total_steps, na.rm = TRUE)
mean_steps
median_steps
```{r daily-pattern}
average_daily_pattern <- data %>%
  group_by(interval) %>%
  summarize(avg_steps = mean(steps, na.rm = TRUE))
plot(average_daily_pattern$interval, average_daily_pattern$avg_steps, type = "l",
     main = "Average Daily Activity Pattern", xlab = "5-Minute Interval", ylab = "Average Steps")
max_interval <- average_daily_pattern[which.max(average_daily_pattern$avg_steps), ]
max_interval
```{r imputing-missing-values}
missing_values <- sum(is.na(data$steps))
missing_values

imputed_data <- data
imputed_data$steps <- ifelse(is.na(imputed_data$steps),
                             average_daily_pattern$avg_steps[match(imputed_data$interval, average_daily_pattern$interval)],
                             imputed_data$steps)

sum(is.na(imputed_data$steps))

imputed_steps_per_day <- imputed_data %>%
  group_by(date) %>%
  summarize(total_steps = sum(steps))

hist(imputed_steps_per_day$total_steps, breaks = 20, main = "Total Steps per Day (Imputed)", xlab = "Total Steps", ylab = "Frequency")

mean_imputed_steps <- mean(imputed_steps_per_day$total_steps)
median_imputed_steps <- median(imputed_steps_per_day$total_steps)

mean_imputed_steps
median_imputed_steps
```{r weekdays-weekends}
imputed_data$date <- as.Date(imputed_data$date)
imputed_data$day_type <- ifelse(weekdays(imputed_data$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")

weekday_vs_weekend <- imputed_data %>%
  group_by(interval, day_type) %>%
  summarize(avg_steps = mean(steps))

library(ggplot2)
ggplot(weekday_vs_weekend, aes(x = interval, y = avg_steps, color = day_type)) +
  geom_line() +
  facet_wrap(~day_type, ncol = 1) +
  labs(title = "Activity Patterns: Weekdays vs Weekends",
       x = "5-Minute Interval", y = "Average Steps")
